---
title: "bfastSpatial Peru"
author: "Jan Verbesselt and Loïc Dutrieux"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: slidy_presentation
---

# Getting started!

After bfastSpatial is installed:

  * define path for reading and saving files
  * set directories

```{r, eval=FALSE}
path <- '~/locationofyourscript'
```

```{r, echo=FALSE, message=FALSE,results='hide'}
library(bfastSpatial)
path <- "~/Git/ChangeMonitor-WUR/Talks/bfastSpatial-2016"
```


```{r, message=TRUE, echo=TRUE, error=FALSE, results='hide'}
library(bfastSpatial)
tmpDir <- rasterOptions()$tmpdir
# set the path the to the location of your script
inDir <- file.path(path, 'data')
# StepDir is where we store intermediary outputs
stepDir <- file.path(inDir, 'datastep')
# directory for landsat data
landsatDir <- file.path(stepDir, 'landsat')
# it is where individual NDVI layers will be stored before being stacked
# ndviDir is a subdirectory of stepDir
ndviDir <- file.path(stepDir, 'ndvi')
# OutDir is where the output will be stored
outDir <- file.path(inDir, 'out')
```

# Check if all directories exist

```{r, echo=TRUE}
## check if dir exist
for (i in c(stepDir, ndviDir, outDir, landsatDir)) {
  print(dir.exists(i))
}
```

Download the archive with the data (~100MB)!

```{r}
if (!file.exists(file.path(inDir, 'hungaryData.tar.gz'))) {
  download.file('http://www.geo-informatie.nl/dutri001/hungaryData.tar.gz',
              destfile = file.path(inDir, 'hungaryData.tar.gz'))
  # Alternative download: https://googledrive.com/host/0B6NK5CdpGEa7ajFuQ1dBZHUwMW8
  # Unpack the data
  untar(file.path(inDir, 'hungaryData.tar.gz'), exdir = landsatDir)
}
```

# Preprocessing the data: I

Check files:

```{r}
# Check that we have the right data in the inDir folder
head(getSceneinfo(list.files(landsatDir)))
```

>  Note: the lines below take some time!

# Preprocessing the data: II

```{r}
if (!file.exists(file.path(inDir, 'ndvi_stack.grd'))) {
  # unzip individual file, use the cloudmask, create ndvi (if this does not exists)
  processLandsatBatch(x = landsatDir, outdir = ndviDir, srdir = tmpDir,
                    delete = TRUE, mask = 'fmask', vi = 'ndvi')

  # Make temporal ndvi stack
  ndviStack <- timeStack(x = ndviDir, pattern = glob2rx('*.grd'),
                       filename = file.path(inDir, 'ndvi_stack.grd'),
                       datatype = 'INT2S')
} else {
  ndviStack <- brick(file.path(inDir, 'ndvi_stack.grd'))
}
```

# Derive the overal median value

```{r, fig.height = 4, fig.width = 4}
if (!file.exists(fn <- file.path(outDir, 'medianVI.grd')))
  # median values for all layers
  medVI <- summaryBrick(ndviStack, fun=median, na.rm=TRUE, 
              filename = fn) else {
  medVI <- brick(fn)
}
plot(medVI/10000)
```

# BfmPixel I

Run BfastMonitor for one pixel time series, by clicking on the map:

```{r, fig.height = 4, fig.width = 4}
# Plot a cloud free recent NDVI layer
names(ndviStack)[394]
plot(ndviStack, 394)
```

# BfmPixel II

```{r, eval=FALSE}
# Call bfmPixel in interactive mode
# to test settings before running it on the brick
bfmPixel(x = ndviStack, start = 2014,
        formula=response~harmon, order=1, history = c(2008, 7),
         interactive = TRUE,
         plot = TRUE)
```

```{r, echo = FALSE, fig.height = 4}
bfmPixel(x = ndviStack, cell=17934, start = 2014,
        formula=response~harmon, order=1, history = c(2008, 7),
         plot = TRUE)
```

# bfastSpatial I

```{r}
# Run bfastmonitor (that part takes a long time)
if (!file.exists(fn <- file.path(outDir, 'bfm_2015_harmon_1.grd'))) {
  # make sure the filename is correct and the folder exists
  bfm <- bfmSpatial(x = ndviStack, pptype = 'irregular',
                  start = 2014, history = c(2008, 7),
                  formula = response ~ harmon,
                  order = 1, mc.cores = 3,
                  filename = file.path(outDir, 'bfm_2015_harmon_1.grd')
                  )
} else {
  bfm <- brick(fn)
}
```

# bfastSpatial: break time!

```{r, fig.height = 3.5, fig.width = 8}
# extract change raster
change <- raster(bfm, 1)
# look at the timing of change
months <- changeMonth(change)
# set up labels and colourmap for months
monthlabs <- c("jan", "feb", "mar", "apr", "may", "jun",
               "jul", "aug", "sep", "oct", "nov", "dec")
cols <- rainbow(12)
plot(months, col=cols, breaks=c(1:12), legend=FALSE)
legend("bottomright", legend=monthlabs, cex=0.5, fill=cols, ncol=2)
```

# bfastSpatial: Magnitude of the break

```{r, fig.height = 4, fig.width = 8}
# extract magnitude of the raster and scale values between 0-1.
magn <- raster(bfm, 2) / 10000
# make a version showing only breakpoing pixels
magn_bkp <- magn
magn_bkp[is.na(change)] <- NA
opar <- par(mfrow=c(1, 2))
plot(magn_bkp, main="Magnitude of a breakpoint")
plot(magn, main="Magnitude: all pixels")
par(opar)
```

# bfastSpatial: Set a treshold

```{r, fig.height = 4, fig.width = 4, message=FALSE}
# extract and rescale magnitude and apply a threshold
magn_bkp_thresh <- magn_bkp
magn_bkp_thresh[magn_bkp > -0.05] <- NA
# compare
library(colorspace) ## package for nice colors
cols <- heat_hcl(12)
plot(magn_bkp_thresh, main="magnitude < -0.05", col = cols)
```

# bfastSpatial: check pixel time series

```{r}
# one pixel time
bfmPixel(x = ndviStack, start = 2014, cell=11521,
        formula=response~harmon, order=1, history = c(2008, 7),
         plot = TRUE)
```



